---
title: "Data Inventory Script"
author: "L. Patterson"
date: "August, 2019"
output: 
  html_document:
    toc: true
    theme: united
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<style type="text/css">
  body{ /* Normal */
    font-family: Calibri;
    font-size: 15pt;
  }
</style>

<br /><br />

# Read in Data

The following script can be run in R to generate the tables and graphs observed in the data inventory website: `LINK GOES HERE`
 

```{r libs, message=FALSE, warning=FALSE}
#Load in libraries
library(dplyr); library(magrittr); library(tidyr);
library(plotly); library(ggplot2); library(tidyverse); library(readxl); #in tidyverse
library(visNetwork);

#removes anything stored in memory
rm(list=ls())

#Set up directories. swd is the main directory where the spreadsheets are located. You may have multiple inventories in different folders. Here, we have inventories for the federal government and three states. 

#main directory  *** USER INPUT REQUIRED HERE ***
swd = "C:\\Users\\DIRECTORY PATH TO INVENTORY FOLDER\\"
folderName <- "federal\\";   #federal, ca, tx, nc
fileName <- "Federal";       #Federal, CA, TX, NC

#Load in the individual tabs on the worksheets
#paste0 concatenates the working directory, folder, filenames, etc. to load the excel spreadsheet. 
orgNodes <- read_excel(paste0(swd,folderName,fileName,"_DataInventory.xlsx"), sheet="orgNodes")

orgEdges <- read_excel(paste0(swd,folderName,fileName,"_DataInventory.xlsx"), sheet="orgEdges") #this is optional to create in excel or in R

platformNodes <- read_excel(paste0(swd,folderName,fileName,"_DataInventory.xlsx"), sheet="platformNodes")

dataNodes <- read_excel(paste0(swd,folderName,fileName,"_DataInventory.xlsx"), sheet="dataNodes")

dataNodeTypes <- read_excel(paste0(swd,folderName,fileName,"_DataInventory.xlsx"), sheet="dataNodeTypes");  
  #remove notes and extraneous columns
  dataNodeTypes <- dataNodeTypes[,c(1:9)]

dataEdges <- read_excel(paste0(swd,folderName,fileName,"_DataInventory.xlsx"), sheet="dataEdges") # this is optional to create in excel or in R

```

<br />

# Create Edge Tables
The `orgEdges` and `dataEdges` worksheets can be manually create or created in R. The following scripts will create the edge files if they have not already been generated. 

```{r createEdges}
#########################################################################################
#Create orgEdges Table 
#########################################################################################
#create data frame
lastRow <- dim(orgNodes)[1]
orgEdges2 <- as.data.frame(matrix(nrow=lastRow-1, ncol=2));    
	colnames(orgEdges2) <- c("from","to");

#populate first row
orgEdges2$from <- orgNodes[c(2:lastRow),]$entityIDAbove
orgEdges2$to <- orgNodes[c(2:lastRow),]$entityID

#Sort org edges regardless
orgEdges <- orgEdges %>% arrange(from,to) %>% as.data.frame()
head(orgEdges)
```

<br />

```{r createDataEdges}
#########################################################################################
#Create dataEdges Table
#########################################################################################
#create data frame
dataEdges <- as.data.frame(matrix(nrow=dim(dataNodeTypes)[1], ncol=2));    colnames(dataEdges) <- c("from","to");
#populate first row
dataEdges$from <- dataNodeTypes$platformID;
#populate the to column 
dataEdges$to <- dataNodeTypes$dataID
#remove duplicates if any and sort dataEdges
dataEdges <- dataEdges %>% distinct(from,to) %>% arrange(from, to) %>% as.data.frame();
head(dataEdges)
```

<br />

# Create Organizational Network

The following script will create a simple heirarchy of the organization. The organization can be changed to display those with a specific purpose by sub-setting. An example is provided below.
```{r createheirarch, message=FALSE, warning=FALSE}
# create nodes in format
nodes <- as.data.frame(matrix(nrow=dim(orgNodes)[1],ncol=0))
  nodes$id <- orgNodes$entityID;
  nodes$label <- gsub(" ", "\n", orgNodes$entity);
  nodes$title <- nodes$label;
  nodes$operational <- orgNodes$operational;
  nodes$decision <- orgNodes$decision;
  nodes$regulatory <- orgNodes$regulatory;
  nodes$research <- orgNodes$research;

  nodes$shape <- "square";
  nodes$color <- "darkgray";
  nodes$shadow <- FALSE;
visNetwork(nodes, orgEdges, main="Organizational Layout", width="800px", height="400px") %>% visHierarchicalLayout()

```

<br />

```{r createheirarchSubset}
#Create network showing regulatory data
nodes.Reg <- subset(nodes, regulatory=="Yes");
edges.Reg <- orgEdges[orgEdges$from %in% nodes.Reg$id,]
visNetwork(nodes.Reg, edges.Reg, main="Regulatory Organizations") %>% visHierarchicalLayout()
```


# Create network for data platforms

This code will create a file used to draw the platform network. 

```{r createdataframe}
####################################################################################################################################
#combine platformNodes and dataNodes into a single file for javascript
####################################################################################################################################
#create dataframe
platformNodes$group <- "hub";
comboNodes <- platformNodes %>% as.data.frame();
  start <- dim(comboNodes)[1]+1;     
  end <- dim(dataNodes)[1]+start-1; 
comboNodes <- rbind(comboNodes, comboNodes[c(1:dim(dataNodes)[1]),]);
comboNodes[c(start:end),] <- NA

#create new ids
comboIDs <- c(platformNodes$platformID,dataNodes$dataID)
comboNodes$platformID <- comboIDs;
  
#create new platform names
comboNames <- c(platformNodes$platform, dataNodes$dataGroup)
comboNodes$platform <- comboNames;

#create new web labels
comboLabel <- c(platformNodes$webLabel, dataNodes$dataCategory)
comboNodes$webLabel <- comboLabel;

#create new groups
comboGroup <- c(platformNodes$group, dataNodes$dataCategory)
comboNodes$group <- comboGroup;

#Count number of occurrences --- sets the size of the nodes
comboNodes$count = 5;
toCount = table(dataEdges$to) %>% as.data.frame(); 
  #toCount;
comboNodes <- merge(comboNodes, toCount, by.x="platformID", by.y="Var1", all = TRUE)
  comboNodes$count = ifelse(comboNodes$group != "hub" & is.na(comboNodes$Freq)==TRUE, 0, comboNodes$count)
  comboNodes$count = ifelse(is.na(comboNodes$Freq)==TRUE, comboNodes$count, comboNodes$Freq)
#comboNodes$count = comboNodes$count+5;
comboNodes <- comboNodes %>% select(-Freq)
#head(comboNodes);
```

<br />

This code will create a file used to draw the platform network. 
```{r drawnetwork, message=FALSE, warning=FALSE}
# create nodes in format
nodesP <- as.data.frame(matrix(nrow=dim(comboNodes)[1],ncol=0))
  nodesP$id <- comboNodes$platformID;
  #nodesP$label <- gsub(" ", "\n", comboNodes$webLabel);
  nodesP$title = ifelse(comboNodes$group=="hub", paste0("<p style='color: black';><strong>", comboNodes$webLabel, "</strong></p>"), 
                                                        paste0("<p style='color: black';><strong>", comboNodes$platform, "</strong></p>"))
    nodesP[,c(3:16)] <- comboNodes[,c(10:23)]

  nodesP$shape <- ifelse(nodesP$group=="hub","square","triangle");
  nodesP$color <- ifelse(nodesP$group=="hub","darkgray","red");
    nodesP$color <- ifelse(nodesP$group=="Quality","purple",nodesP$color);
    nodesP$color <- ifelse(nodesP$group=="Quantity","lightskyblue",nodesP$color);
    nodesP$color <- ifelse(nodesP$group=="Use","seagreen",nodesP$color);
  nodesP$size <- nodesP$count*3.5;
  
#create edges and format
  edgesD <- merge(dataEdges, nodesP[,c("id","color")], by.x="to", by.y="id")
  edgesD$width <- 2
  
#draw network    
visNetwork(nodesP, edgesD, main = "Network of Platforms and Data Types", height="600px", width="600px") 

```

<br />
Similarly, the nodes can be highlighted and the graph redrawn based on your interest. The example below will show how. More complicated parsing of data can be found in the scorecard section below.
```{r subsetdataframe, message=FALSE, warning=FALSE}
#select those hubs that provide glossaries for their data.
unique(nodesP$dataDefinitions)
nodesPHigh <- nodesP;
nodesPHigh$color <- ifelse(nodesPHigh$dataDefinitions=="Yes" & nodesPHigh$group=="hub", "goldenrod", nodesPHigh$color);


#redraw network
visNetwork(nodesPHigh, edgesD, main = "Highlight Platforms with Data Glossaries", height="600px", width="600px") %>% visLayout(randomSeed = 12) # to have always the same network     
```

<br />

# Create Heatmap

Below is the code to create a heatmap table. The table will then be visualized.
``` {r heatmap}
####################################################################################################################################
#Create heatmap table
####################################################################################################################################
#create data frame
heatmap <- as.data.frame(matrix(nrow=dim(platformNodes)[1], ncol=(3+dim(dataNodes)[1])));

#create column names from dataNodes
heatmapColNames <- c(str_replace_all(dataNodes$dataGroup, " ", ""))
colnames(heatmap) <- c("platformID", "entityID", "platform", heatmapColNames)
#colnames(heatmap) <- c("platformID", "entityID", "platform", "Built",	"Natural", "Quality", "Regulatory", "ET", "ExtremeEvents", "GlacialandSnow",
#                       "Groundwater", "Meteorology", "Precipitation", "Reservoir", "Soil", "SurfaceWater", "Hydropower", "Irrigation", "Use", "Utilities", "ManagementPlans")

#fill first three columns
heatmap$platformID <- platformNodes$platformID;
heatmap$entityID <- platformNodes$entityID;
heatmap$platform <- platformNodes$heatmapLabel;

#loop through and subset dataEdges based on platform ID
for(i in 1:dim(platformNodes)[1]){
  zt.edges <- subset(dataEdges, from==platformNodes$platformID[i]) %>% as.data.frame();
  
  #if the value exists, then fill in with a 1 [row, col]
  heatmap[i,4] <- ifelse(dim(subset(zt.edges, to=="d01"))[1]==1,1,0);          heatmap[i,5] <- ifelse(dim(subset(zt.edges, to=="d02"))[1]==1,1,0);
  heatmap[i,6] <- ifelse(dim(subset(zt.edges, to=="d03"))[1]==1,3,0);          heatmap[i,7] <- ifelse(dim(subset(zt.edges, to=="d04"))[1]==1,3,0);
  heatmap[i,8] <- ifelse(dim(subset(zt.edges, to=="d05"))[1]==1,2,0);          heatmap[i,9] <- ifelse(dim(subset(zt.edges, to=="d06"))[1]==1,2,0);
  heatmap[i,10] <- ifelse(dim(subset(zt.edges, to=="d07"))[1]==1,2,0);         heatmap[i,11] <- ifelse(dim(subset(zt.edges, to=="d08"))[1]==1,2,0);
  heatmap[i,12] <- ifelse(dim(subset(zt.edges, to=="d09"))[1]==1,2,0);         heatmap[i,13] <- ifelse(dim(subset(zt.edges, to=="d10"))[1]==1,2,0);
  heatmap[i,14] <- ifelse(dim(subset(zt.edges, to=="d11"))[1]==1,2,0);         heatmap[i,15] <- ifelse(dim(subset(zt.edges, to=="d12"))[1]==1,2,0);
  heatmap[i,16] <- ifelse(dim(subset(zt.edges, to=="d13"))[1]==1,2,0);         heatmap[i,17] <- ifelse(dim(subset(zt.edges, to=="d14"))[1]==1,4,0);
  heatmap[i,18] <- ifelse(dim(subset(zt.edges, to=="d15"))[1]==1,4,0);         heatmap[i,19] <- ifelse(dim(subset(zt.edges, to=="d16"))[1]==1,4,0);
  heatmap[i,20] <- ifelse(dim(subset(zt.edges, to=="d17"))[1]==1,4,0);         heatmap[i,21] <- ifelse(dim(subset(zt.edges, to=="d18"))[1]==1,4,0);
}

#sort by entityID and then platform
heatmap <- heatmap %>% arrange(platform)
head(heatmap)
```

<br />

Next, we will draw the heatmap using the plotly library.
``` {r heatmapDraw, message=FALSE, warning=FALSE}
#create color scheme
cols <- c("0" = "lightgray", "1" = "red", "2" = "lightskyblue", "3" = "purple", "4" = "seagreen")

data=as.matrix(t(heatmap[,c(4:21)]))
plot_ly(x=heatmap$platform, y=colnames(heatmap[,c(4:21)]), z = data, type = "heatmap", color=c(0,1,2,3,4), colors=cols, showscale = FALSE, xgap=1, ygap=1) %>%
    layout(
        title = "<b>Heat map showing who provides what data</b>",
        width=800,
        xaxis=list(tickfont = list(size = 10), tickangle = 45),
        margin = list(l = 60, r = 10, b = 100, t = 30, pad = 4),
        showlegend = FALSE) 
```

<br />

The following script summarizes the rows - showing the number of entities collecting each type of water data.
``` {r heatmap_rowsum, message=FALSE, warning=FALSE}
#summarize each of the columns
#make presence a 1 (not 2, 3, or 4)
heatmap2 <- heatmap
heatmap2[heatmap2>=2] <- 1
  
nTypes <- heatmap2 %>%  summarise_if(is.numeric, sum, na.rm = TRUE)
nTypes <- t(nTypes) %>% as.data.frame(); 
#set row names to column
colnames(nTypes) <- c("count")
nTypes <- tibble::rownames_to_column(nTypes, "name")

#graph - unfortunately plotly automatically re-orders charge alphabetically. It is a known bug in the program.
plot_ly(
  x = nTypes$name,
  y = nTypes$count,
  #color = nTypes$color,
  marker = list(color = c("red","red","purple","purple","lightskyblue","lightskyblue","lightskyblue","lightskyblue","lightskyblue",
                  "lightskyblue","lightskyblue","lightskyblue","lightskyblue","seagreen", "seagreen", "seagreen", "seagreen", "seagreen")),
  type = "bar", 
  orientation = "vertical"
) %>%
  layout(width=800,
        title = "<b>What types of data are most often collected?</b>",
        yaxis = list(title = 'Number of Entities Collecting Data'), 
        xaxis=list(tickfont = list(size = 10), tickangle = 45),
        margin = list(l = 50, r = 10, b = 90, t = 30, pad = 4),
        showlegend = FALSE)
````

<br />

The following script summarizes the rows - showing the number of entities collecting each type of water data.
``` {r heatmap_colSum, message=FALSE, warning=FALSE}
#summarize each of the rows
nEntities <- heatmap %>% mutate(count = rowSums(.[4:21]))

#graph - unfortunately plotly automatically re-orders charge alphabetically. It is a known bug in the program.
plot_ly(
  x = nEntities$platform,
  y = nEntities$count,
  #color = nTypes$color,
  marker = list(color = c("navy")),
  type = "bar", 
  orientation = "vertical"
) %>%
  layout(width = 800,
        title = "<b>Which platforms provide the greatest variety of water-related data?</b>",
        yaxis = list(title = 'Number of Data Types Collected'), 
        xaxis=list(tickfont = list(size = 10), tickangle = 45),
        margin = list(l = 50, r = 30, b = 90, t = 30, pad = 4),
        showlegend = FALSE)

```


<br />

# Creating the score card metrics
The following scripts will create scores based on discoverability, accessibility, and usability characteristics of the data platform. Descriptions and reasoning behind the scores can be found in the white paper <strong><a herf=>Inventory of Federal and State Water Data</a</strong>.

## Discoverability - how easy was it to find data?
``` {r easeDescover}
#create score for ease of discovery ---------------------------------------------------------------------------
scoreNodes <- platformNodes;
#make all lowercase 
scoreNodes$easeDiscover <- tolower(scoreNodes$easeDiscover)
#find unique
unique(scoreNodes$easeDiscover); #should be no, low, medium, or high
#create score
scoreNodes$scoreEaseDiscover = NA;
  scoreNodes$scoreEaseDiscover = ifelse(scoreNodes$easeDiscover == "no" | scoreNodes$easeDiscover=="unknown" | scoreNodes$easeDiscover=="none", 0, scoreNodes$scoreEaseDiscover);
  scoreNodes$scoreEaseDiscover = ifelse(scoreNodes$easeDiscover == "low", 1, scoreNodes$scoreEaseDiscover);
  scoreNodes$scoreEaseDiscover = ifelse(scoreNodes$easeDiscover == "medium", 2, scoreNodes$scoreEaseDiscover);
  scoreNodes$scoreEaseDiscover = ifelse(scoreNodes$easeDiscover == "high", 3, scoreNodes$scoreEaseDiscover);
#check to see all values scored
table(scoreNodes$scoreEaseDiscover, useNA = "ifany");  
```

<br />

## Discoverability - how were data found?
``` {r methodDescover}
#create score for method to find data ---------------------------------------------------------------------------
#make all lowercase
scoreNodes$methodDiscover <- tolower(scoreNodes$methodDiscover);
#find unique
unique(scoreNodes$methodDiscover);  #should be some version of unknoqne, website, map, catalog
#create score
scoreNodes$scoreMethodDiscover = 0;
#loop through and separate by "; " to find all the values. We will use the mean score.
for (i in 1:dim(scoreNodes)[1]){
  foo = as.list(strsplit(as.character(scoreNodes$methodDiscover[i]),"; "));
  foo = as.data.frame(foo); colnames(foo) = "value";
  foo$score = 0;
  foo$score = ifelse(foo$value=="website",1, foo$score);
  foo$score = ifelse(foo$value=="map",2, foo$score);
  foo$score = ifelse(foo$value=="catalog",2, foo$score);
  
  scoreNodes$scoreMethodDiscover[i] = round(mean(foo$score),2);  
}
#check to make sure all nodes scored
table(scoreNodes$scoreMethodDiscover, useNA = "ifany"); 
```

<br />

## Accessibility - how easy was it to access data?
``` {r easeAccess}
# create score for ease of access  ---------------------------------------------------------------------------
#make all lowercase
scoreNodes$easeAccess <- tolower(scoreNodes$easeAccess);
#find unique values
unique(scoreNodes$easeAccess);  #yes, no, permission required, software required, training required, registration required
#create score  
scoreNodes$scoreEaseAccess = NA;
  scoreNodes$scoreEaseAccess <- ifelse(scoreNodes$easeAccess=="unknown" | scoreNodes$easeAccess=="no", 0, scoreNodes$scoreEaseAccess)
  scoreNodes$scoreEaseAccess <- ifelse(scoreNodes$easeAccess=="software required" | scoreNodes$easeAccess=="training required" | 
                                           scoreNodes$easeAccess == "permission required", 1, scoreNodes$scoreEaseAccess)
  scoreNodes$scoreEaseAccess <- ifelse(scoreNodes$easeAccess=="registration required", 3, scoreNodes$scoreEaseAccess)  #assumes permission granted
  scoreNodes$scoreEaseAccess <- ifelse(scoreNodes$easeAccess=="yes", 3, scoreNodes$scoreEaseAccess)
#Check to make sure all platforms are scored
table(scoreNodes$scoreEaseAccess, useNA = "ifany"); 
```


<br />

## Accessibility - how were data obtained?
``` {r methodAccess}
# create score for method of access ---------------------------------------------------------------------------
#make all lower case
scoreNodes$methodAccess <- tolower(scoreNodes$methodAccess)
#find unique values
unique(scoreNodes$methodAccess)
#create score
scoreNodes$scoreMethodAccess = NA;
for (i in 1:dim(scoreNodes)[1]){
  foo = as.list(strsplit(as.character(scoreNodes$methodAccess[i]),"; "));
  foo = as.data.frame(foo); colnames(foo) = "value";
    foo$score = NA;
    foo$score = ifelse(foo$value=="unknown" | foo$value=="none",0, foo$score);
    foo$score = ifelse(foo$value=="link to source" | foo$value=="copy and paste",1, foo$score);
    foo$score = ifelse(foo$value=="individual export",2, foo$score);
    foo$score = ifelse(foo$value=="batch export" | foo$value=="full export",3, foo$score);
    foo$score = ifelse(foo$value=="ftp", 4, foo$score);
    foo$score = ifelse(foo$value=="web services",5, foo$score);
  
  #take the maximum score
  scoreNodes$scoreMethodAccess[i] = round(sum(foo$score)/dim(foo)[1],2);  
}
#check to make sure all platforms are scored
table(scoreNodes$scoreMethodAccess, useNA = "ifany") 
#subset(scoreNodes, is.na(scoreMethodAccess)==TRUE) %>% as.data.frame() #make sure not missing any


```


<br />

## Interoperability - scoring data file formats - machine readable?
``` {r fileFormat}
#create score for file type  ---------------------------------------------------------------------------
#make all lower case
scoreNodes$fileFormat <- tolower(scoreNodes$fileFormat);
#find unique values
unique(scoreNodes$fileFormat)
#create score
scoreNodes$scoreFileFormat = NA
#loop through and separate by ";" - will then take the mean of the file formats
for (i in 1:dim(scoreNodes)[1]){
  foo = as.list(strsplit(as.character(scoreNodes$fileFormat[i]),"; "));
    foo = as.data.frame(foo); colnames(foo) = "value";
  foo$score = NA;
  foo$score = ifelse(foo$value=="unknown" | foo$value=="not applicable" | foo$value=="none",0, foo$score);
  foo$score = ifelse(foo$value=="pdf" | foo$value=="png" | foo$value=="image" | foo$value=="jpg" | foo$value=="zip",1, foo$score);   #not sure where to put zip
  
  foo$score = ifelse(foo$value=="access" | foo$value=="excel" | foo$value=="raster" | foo$value=="shapefile",3, foo$score);
  foo$score = ifelse(foo$value=="word" | foo$value=="geodatabase" | foo$value=="kmz" | foo$value=="software specific",3, foo$score);
  
  foo$score = ifelse(foo$value=="csv" | foo$value=="json" | foo$value=="txt" | foo$value == "xml",5, foo$score);
  foo$score = ifelse(foo$value=="ascii" | foo$value=="netcdf" | foo$value=="html" | foo$value == "hdf5",5, foo$score);
  foo$score = ifelse(foo$value=="kml" | foo$value=="geojson" | foo$value == "wml2" | foo$value == "rdf" | foo$value == "tsv",5, foo$score);
  
  scoreNodes$scoreFileFormat[i] = round(max(foo$score, is.na=FALSE),2);  
}
#check to make sure all platforms are scored
table(scoreNodes$scoreFileFormat, useNA = "ifany") 
#subset(scoreNodes, is.na(scoreFileFormat)==TRUE)  #if find a bug (such as a last ; with nothing after it - go into original file and fix)
```


<br />

## Interoperability - scoring metadata file format - machine readable?
``` {r metdataFormat}
#create score for metadata -------------------------------------------------------------------------------
#make all lower case
scoreNodes$metadataFormat <- tolower(scoreNodes$metadataFormat)
#find unique values
unique(scoreNodes$metadataFormat)
#create values for machine read and for non-machine read
no.machineread <- c("in data", "pdf")
prop.machineread <- c("word", "excel")
yes.machineread <- c("xml","html","json","txt", "csv")
#score
scoreNodes$scoreMetadataFormat <- NA

for (i in 1:dim(scoreNodes)[1]){
  foo = as.list(strsplit(as.character(scoreNodes$metadataFormat[i]),"; "));
  foo = as.data.frame(foo); colnames(foo) = "value";
  foo$score = NA;
  foo$score = ifelse(foo$value=="no" | foo$value=="broken link",0, foo$score);
  foo$score = ifelse(foo$value %in% no.machineread, 1, foo$score);
  foo$score = ifelse(foo$value %in% prop.machineread, 2, foo$score);
  foo$score = ifelse(foo$value %in% yes.machineread, 3, foo$score);

  scoreNodes$scoreMetadataFormat[i] = round(mean(foo$score, is.na=FALSE),2);  
}
scoreNodes$scoreMetadataFormat <- ifelse(scoreNodes$metadataFormat == "unknown" & scoreNodes$metadata=="No", 0, scoreNodes$scoreMetadataFormat)
scoreNodes$scoreMetadataFormat <- ifelse(scoreNodes$metadataFormat == "unknown" & scoreNodes$metadata=="Yes", 1, scoreNodes$scoreMetadataFormat)
scoreNodes$scoreMetadataFormat <- ifelse(scoreNodes$metadataFormat == "unknown" & scoreNodes$metadata=="Unknown", 1, scoreNodes$scoreMetadataFormat)

#check to make sure all platforms are scored
table(scoreNodes$scoreMetadataFormat, useNA = "ifany") 
#subset(scoreNodes, is.na(scoreMetadataFormat)==TRUE)  #if find a bug (such as a last ; with nothing after it - go into original file and fix)
```

<br />

## Interoperability - is there a glossary for the data?
``` {r dataDefinitions}
#create score for metadata -------------------------------------------------------------------------------
#make all lower case
scoreNodes$dataDefinitions <- tolower(scoreNodes$dataDefinitions)
#find unique values
unique(scoreNodes$dataDefinitions)
#score
scoreNodes$scoreDataDefinitions <- NA
scoreNodes$scoreDataDefinitions <- ifelse(scoreNodes$dataDefinitions == "no" | scoreNodes$dataDefinitions=="unknown", 0, scoreNodes$scoreDataDefinitions)
  scoreNodes$scoreDataDefinitions <- ifelse(scoreNodes$dataDefinitions == "some", 1, scoreNodes$scoreDataDefinitions)
  scoreNodes$scoreDataDefinitions <- ifelse(scoreNodes$dataDefinitions == "yes", 2, scoreNodes$scoreDataDefinitions)
table(scoreNodes$scoreDataDefinitions, useNA = "ifany") 
```

<br /> 

## Interoperability - what types of metadata are provided?
```{r metdataAttributes}
#create score for metadata -------------------------------------------------------------------------------
#metadata will be composed of four attributes - not the yes/no present question
#make all lower case
scoreNodes$metaAttributes <- tolower(scoreNodes$metaAttributes)
#find unique values
unique(scoreNodes$metaAttributes)
#create score
scoreNodes$scoreMetaAttributes = NA
#loop through and separate by ";" - will then take the mean of the file formats
for (i in 1:dim(scoreNodes)[1]){
  foo = as.list(strsplit(as.character(scoreNodes$metaAttributes[i]),"; "));
  foo = as.data.frame(foo); colnames(foo) = "value";
  foo$score = NA;
  foo$score = ifelse(foo$value=="unknown" | foo$value=="no",0, foo$score);
  foo$score = ifelse(foo$value=="administrative" | foo$value=="structural" | foo$value=="descriptive", 1, foo$score);

  scoreNodes$scoreMetaAttributes[i] = round(sum(foo$score, is.na=FALSE),2);  
}

#check to make sure platforms scored
table(scoreNodes$scoreMetaAttributes, useNA = "ifany") 
#subset(scoreNodes, is.na(scoreMetaAttributes)==TRUE)  #if find a bug (such as a last ; with nothing after it - go into original file and fix)
```

<br />

## Interoperability - are metadata standards provided?
```{r metadataStandards}
#create score for metadata -------------------------------------------------------------------------------
#metadata will be composed of four attributes - not the yes/no present question
#make all lower case
scoreNodes$metadataStandards <- tolower(scoreNodes$metadataStandards)
#find unique values
unique(scoreNodes$metadataStandards)
#create score
scoreNodes$scoreMetadataStandards <- NA
  scoreNodes$scoreMetadataStandards <- ifelse(scoreNodes$metadataStandards == "no" | scoreNodes$metadataStandards=="unknown", 0, scoreNodes$scoreMetadataStandards)
  scoreNodes$scoreMetadataStandards <- ifelse(scoreNodes$metadataStandards == "some", 1, scoreNodes$scoreMetadataStandards)
  scoreNodes$scoreMetadataStandards <- ifelse(scoreNodes$metadataStandards == "yes", 2, scoreNodes$scoreMetadataStandards)

#check to make sure platforms scored
table(scoreNodes$scoreMetadataStandards, useNA = "ifany") #should have 18 for data nodes
```


<br />

## Interoperability - how often are data being updated? 

```{r timeliness}
#create score for timeliness -------------------------------------------------------------------------------  
#make all lower case
scoreNodes$timeliness <- tolower(scoreNodes$timeliness)
#find unique values
unique(scoreNodes$timeliness)
#create score
scoreNodes$scoreTimeliness <- NA
  scoreNodes$scoreTimeliness = ifelse(scoreNodes$timeliness=="unknown", 0, scoreNodes$scoreTimeliness);
  scoreNodes$scoreTimeliness = ifelse(scoreNodes$timeliness=="irregular" | scoreNodes$timeliness=="varies", 1, scoreNodes$scoreTimeliness);
  scoreNodes$scoreTimeliness = ifelse(scoreNodes$timeliness=="yearly or higher", 2, scoreNodes$scoreTimeliness);
  scoreNodes$scoreTimeliness = ifelse(scoreNodes$timeliness=="monthly or higher" | scoreNodes$timeliness=="weekly or higher", 3, scoreNodes$scoreTimeliness);
  scoreNodes$scoreTimeliness = ifelse(scoreNodes$timeliness=="daily or higher", 4, scoreNodes$scoreTimeliness);

#check to make sure platforms scored
table(scoreNodes$scoreTimeliness, useNA = "ifany"); #should have 18 for datanodes
#subset(scoreNodes, is.na(timeliness)==TRUE)

```


<br />

## Interoperability - how much of the data are made available?
```{r POR}
#create score for length of record -------------------------------------------------------------------------------  
#make all lower case
scoreNodes$lengthAvailable <- tolower(scoreNodes$lengthAvailable)
#find unique values
unique(scoreNodes$lengthAvailable)
#set score
scoreNodes$scoreLengthAvailable = NA;
  scoreNodes$scoreLengthAvailable = ifelse(scoreNodes$lengthAvailable=="unknown",0, scoreNodes$scoreLengthAvailable);
  scoreNodes$scoreLengthAvailable = ifelse(scoreNodes$lengthAvailable=="not applicable" | scoreNodes$lengthAvailable=="varies", 1, scoreNodes$scoreLengthAvailable);
  scoreNodes$scoreLengthAvailable = ifelse(scoreNodes$lengthAvailable=="current only", 2, scoreNodes$scoreLengthAvailable);
  scoreNodes$scoreLengthAvailable = ifelse(scoreNodes$lengthAvailable=="limited record", 3, scoreNodes$scoreLengthAvailable);
  scoreNodes$scoreLengthAvailable = ifelse(scoreNodes$lengthAvailable=="period of record", 4, scoreNodes$scoreLengthAvailable);

#check to make sur eplatforms scored
table(scoreNodes$scoreLengthAvailable, useNA = "ifany"); #should have 18 for data nodes
#subset(scoreNodes, is.na(lengthAvailable)==TRUE)

```


<br /> <br />

# Setting the scores on a scale from 0 to 100 for comparability
```{r normalizeScores}
###############################################################################################################
#total score for discoverable, accessible, usable
scoreNodes$discoverTotal = scoreNodes$scoreEaseDiscover + scoreNodes$scoreMethodDiscover;
scoreNodes$accessTotal = scoreNodes$scoreEaseAccess + scoreNodes$scoreMethodAccess;  
scoreNodes$usableTotal = scoreNodes$scoreFileFormat + scoreNodes$scoreMetadataFormat + scoreNodes$scoreDataDefinitions + 
                         scoreNodes$scoreMetaAttributes + scoreNodes$scoreMetadataStandards + scoreNodes$scoreTimeliness + 
                         scoreNodes$scoreLengthAvailable
scoreNodes$scoreTotal <- scoreNodes$discoverTotal + scoreNodes$accessTotal + scoreNodes$usableTotal;  
  
  
#max score for each variable
easeDiscoverMax = 3;   methodDiscoverMax = 2;     totalDiscoverMax = easeDiscoverMax + methodDiscoverMax;

easeAccessMax = 3;     methodAccessMax = 5;       totalAccessMax = easeAccessMax + methodAccessMax;

fileFormatMax = 5;     metadataFormatMax = 3;     dataDefinitionsMax = 2;    metaAttributesMax = 3;     metadataStandardsMax = 2;  
  timelinessMax = 4;   lengthAvailableMax = 4;    
  totalUseMax = fileFormatMax + metadataFormatMax + dataDefinitionsMax + metaAttributesMax + metadataStandardsMax + timelinessMax + lengthAvailableMax;

totalMax = totalDiscoverMax + totalAccessMax + totalUseMax;
#######################################################################################################################
  
  
#######################################################################################################################
# CONVERT SCORES TO PERCENT
#######################################################################################################################
percent = scoreNodes;
  
percent$scoreEaseDiscover = round(scoreNodes$scoreEaseDiscover/easeDiscoverMax*100,2);
percent$scoreMethodDiscover = round(scoreNodes$scoreMethodDiscover/methodDiscoverMax*100,2);
  
percent$scoreEaseAccess = round(scoreNodes$scoreEaseAccess/easeAccessMax*100,2);
percent$scoreMethodAccess = round(scoreNodes$scoreMethodAccess/methodAccessMax*100,2);
  
percent$scoreFileFormat = round(scoreNodes$scoreFileFormat/fileFormatMax*100,2);
percent$scoreMetadataFormat = round(scoreNodes$scoreMetadataFormat/metadataFormatMax*100,2);
percent$scoreDataDefinitions = round(scoreNodes$scoreDataDefinitions/dataDefinitionsMax*100,2);
percent$scoreMetaAttributes = round(scoreNodes$scoreMetaAttributes/metaAttributesMax*100,2);
percent$scoreMetadataStandards = round(scoreNodes$scoreMetadataStandards/metadataStandardsMax*100,2);

percent$scoreTimeliness = round(scoreNodes$scoreTimeliness/timelinessMax*100,2);
percent$scoreLengthAvailable = round(scoreNodes$scoreLengthAvailable/lengthAvailableMax*100,2);
  
percent$discoverTotal = round(scoreNodes$discoverTotal/totalDiscoverMax*100,2);
percent$accessTotal = round(scoreNodes$accessTotal/totalAccessMax*100,2);
percent$usableTotal = round(scoreNodes$usableTotal/totalUseMax*100,2);
  
percent$scoreTotal = round(scoreNodes$scoreTotal/totalMax*100, 2);
#summary(percent)

```

<br />

## Create a boxplot of the scores by categories of discoverability, accessibility, and usability
```{r boxplotScoresDiscover, message=FALSE, warning=FALSE}
#create a boxplot of discovaerability / findability
plot_ly(y = percent$scoreEaseDiscover, x="Ease Discover", type = "box",  
        boxpoints = "all", jitter = 0.3, pointpos = -1.8,
          marker = list(color = 'navy'),
          line = list(color = 'navy'),
          name = "Ease Discover") %>%
  add_trace(y = percent$scoreMethodDiscover, x= "Method Discover",
          marker = list(color = 'navy'),
          line = list(color = 'navy'),
          name = "Method Discover") %>% 
  add_trace(y = percent$discoverTotal, x="Total Score",
          marker = list(color = 'navy'),
          line = list(color = 'navy'),
          name = "Total Score") %>% 
  layout(xaxis=list(tickfont = list(size = 12)), yaxis=list(title="Score (Percent)"),
        margin = list(l = 50, r = 10, b = 30, t = 30, pad = 4),
        title = "<b>Discoverability Scores</b>",
        showlegend = FALSE)
```


```{r boxplotScoresAccess, message=FALSE, warning=FALSE}
#create a boxplot of accessibility
plot_ly(y = percent$scoreEaseAccess, x="Ease Access", type = "box",  
        boxpoints = "all", jitter = 0.3, pointpos = -1.8,
          marker = list(color = 'navy'),
          line = list(color = 'navy'),
          name = "Ease Access") %>%
  add_trace(y = percent$scoreMethodAccess, x= "Method Access",
          marker = list(color = 'navy'),
          line = list(color = 'navy'),
          name = "Method Access") %>% 
  add_trace(y = percent$accessTotal, x="Total Score",
          marker = list(color = 'navy'),
          line = list(color = 'navy'),
          name = "Total Score") %>% 
  layout(xaxis=list(tickfont = list(size = 12)), yaxis=list(title="Score (Percent)"),
        margin = list(l = 50, r = 10, b = 30, t = 30, pad = 4),
        title = "<b>Accessibility Scores</b>",
        showlegend = FALSE)
```


```{r boxplotScoreInterop, message=FALSE, warning=FALSE}
#create a boxplot of interoperability / usability
plot_ly(y = percent$scoreFileFormat, x="File Format", type = "box",  
        boxpoints = "all", jitter = 0.3, pointpos = -1.8,
          marker = list(color = 'navy'),
          line = list(color = 'navy'),
          name = "File Format") %>%
  add_trace(y = percent$scoreMetadataFormat, x= "Metadata Format",
          marker = list(color = 'navy'),
          line = list(color = 'navy'),
          name = "Metadata Format") %>% 
  add_trace(y = percent$scoreDataDefinitions, x= "Data Definitions",
          marker = list(color = 'navy'),
          line = list(color = 'navy'),
          name = "Data Definitions") %>% 
  add_trace(y = percent$scoreMetaAttributes, x= "Metadata Types",
          marker = list(color = 'navy'),
          line = list(color = 'navy'),
          name = "Metadata Types") %>% 
  add_trace(y = percent$scoreMetadataStandards, x= "Metadata Standards",
          marker = list(color = 'navy'),
          line = list(color = 'navy'),
          name = "Metadata Standards") %>% 
  add_trace(y = percent$scoreTimeliness, x= "Frequency of Updates",
          marker = list(color = 'navy'),
          line = list(color = 'navy'),
          name = "Frequency of Updates") %>% 
  add_trace(y = percent$scoreLengthAvailable, x= "Amount of Data",
          marker = list(color = 'navy'),
          line = list(color = 'navy'),
          name = "Amount of Data") %>% 
  add_trace(y = percent$usableTotal, x="Total Score",
          marker = list(color = 'navy'),
          line = list(color = 'navy'),
          name = "Total Score") %>% 
  layout(xaxis=list(tickfont = list(size = 12)), yaxis=list(title="Score (Percent)"),
        margin = list(l = 50, r = 10, b = 30, t = 30, pad = 4),
        title = "<b>Interoperability Scores</b>",
        showlegend = FALSE)
```

<br />

The files can be saved by `write.csv(percent, paste0(swd,folderName,fileName,"_percent.csv"), row.names=FALSE)`.